---
title: "Homework 2"
author: "Shen Dingtao  3170104764"
output: pdf_document
---
```{r setup, message = F, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(DAAG)
library(MASS)
```
**1. _Loading and cleaning_**

**(a)**
```{r}
ca_pa <- read.csv("data/calif_penn_2011.csv",header=T)
```
**(b)**
```{r}
dim(ca_pa)
```
So the data frame has 11275 rows and 34 columns.

**(c)**
```{r}
colSums(apply(ca_pa,c(1,2),is.na))
```
This command is used to count the number of NA (not a number) values in each column of the data frame.

- The `apply()` function loops through all the elements of the matrix `ca_pa` and applies the `is.na()` function which returns TRUE if the element is not a number and FALSE otherwise. 

- The resulting matrix of TRUE and FALSE values is then given as input to `colSums()` function, which counts the number of TRUE values in each column.

**(d)**
```{r}
new_ca_pa<-na.omit(ca_pa)
```
**(e)**
```{r}
nrow(ca_pa)-nrow(new_ca_pa)
```
The result shows that the `na.omit()` command eliminate 670 rows. 

**(f)** My answer in (c) and (e) are compatible. We can run the following command to verify this:
```{r}
colSums(apply(new_ca_pa,c(1,2),is.na))
```
The values of all columns are all 0, which means that we have successfully purged ca_pa from any row containing NA value.

**2. _This Very New House_ ** 

**(a)**
```{r}
ggplot(data = new_ca_pa) +
geom_point(aes(x = Built_2005_or_later, 
    y = Median_house_value),na.rm = TRUE) +
labs(x = "Percetsge of Houses Built Since 2005",
y = "Median House Value") +
theme_bw() +
theme(legend.title=element_blank())
```
**(b)** With following commands, we make a pair of plots that breaks this out by state, which shows the median house prices against `Built_2005_or_later` in California(state 6) and Pennsylvania(state 42) respectively.

```{r}
d<-data.frame(STATEFP=c(6,42),state=c("California","Pennsylvania"))
ca_pa2<-left_join(new_ca_pa,d,by="STATEFP")
```

```{r message=FALSE}
ggplot(data = ca_pa2) +
geom_point(aes(x = Built_2005_or_later, 
    y = Median_house_value),na.rm = TRUE) +
labs(x = "Percetsge of Houses Built Since 2005",
y = "Median House Value") +facet_wrap(~ state)
```

**3. _Nobody Home_**

**(a)** Add a new column to the dataframe which contains the vacancy rate.
```{r}
ca_pa3 <- ca_pa2 %>%
  mutate(Vacant_Rate=Vacant_units/Total_units)
```
Minimum:
```{r}
min(ca_pa3$Vacant_Rate,na.rm = TRUE)
```
Maximum:
```{r}
max(ca_pa3$Vacant_Rate,na.rm = TRUE)
```
Mean:
```{r}
mean(ca_pa3$Vacant_Rate,na.rm = TRUE)
```
Median:
```{r}
median(ca_pa3$Vacant_Rate,na.rm = TRUE)
```
**(b)** Plot the vacancy rate against median house value:
```{r message=FALSE}
ggplot(data = ca_pa3) +
geom_point(aes(x = Median_house_value, 
    y = Vacant_Rate),na.rm = TRUE) +
labs(x = "Median House Value",
y = "Vacant Rate",title = "Vacant_Rate vs Median_house_value") 
```
**(c)** Plot vacancy rate against median house value separately for California and for Pennsylvania:
```{r}
ggplot(data = ca_pa3) +
geom_point(aes(x = Median_house_value, 
    y = Vacant_Rate,color=state),na.rm = TRUE) +
facet_wrap(~ state)
```

**4.** 

**(a)** The block of code is supposed to calculate the median house value in Alameda Country (country 1 in California). It firstly selects the rows of California, whose STATEFP=6, and then selects the rows of Alameda, whose COUNTRYFP=1 among the selected rows. Finally select median value of these rows' Median house value to get the median house value in Alameda. 

**(b)** We can obtain the same result as the block of code with following command:
```{r}
ca_pa3 %>% filter(STATEFP==6,COUNTYFP==1) %>% {median(.$Median_house_value,na.rm = TRUE)}
```
**(c)** We can obtain the average percentages of housing built since 2005 for Alameda, Santa Clara and Allegheny Counties with following commands:
(i) For Alameda:
```{r}
(Alameda_avg <- ca_pa3 %>% filter(STATEFP==6&COUNTYFP==1 )) %>% {mean(.$Built_2005_or_later,na.rm = TRUE)}
```
(ii) For Santa Clara:
```{r}
(Alameda_avg <- ca_pa3 %>% filter(STATEFP==6&COUNTYFP==85 )) %>% {mean(.$Built_2005_or_later,na.rm = TRUE)}
```
(ii) For Allegheny:
```{r}
(Alameda_avg <- ca_pa3 %>% filter(STATEFP==42&COUNTYFP==3 )) %>% {mean(.$Built_2005_or_later,na.rm = TRUE)}
```

**(d)** the correlation between median house value and the percent of housing built since 2005
(i) In the whole data:
```{r}
cor(ca_pa3$Median_house_value,ca_pa3$Built_2005_or_later)
```
(ii) In all of California
```{r}
ca_pa3 %>% filter(STATEFP==6) %>% 
  {cor(.$Median_house_value,.$Built_2005_or_later)}
```
(iii) In all of Pennsylvania
```{r}
ca_pa3 %>% filter(STATEFP==42) %>% 
  {cor(.$Median_house_value,.$Built_2005_or_later)}
```
(iv) In Alameda County
```{r}
ca_pa3 %>% filter(STATEFP==6,COUNTYFP==1) %>% 
  {cor(.$Median_house_value,.$Built_2005_or_later)}
```
(v) In Santa Clara County
```{r}
ca_pa3 %>% filter(STATEFP==6,COUNTYFP==85) %>% 
  {cor(.$Median_house_value,.$Built_2005_or_later)}
```
(vi) In Allegheny Count
```{r}
ca_pa3 %>% filter(STATEFP==42,COUNTYFP==3) %>% 
  {cor(.$Median_house_value,.$Built_2005_or_later)}
```
**(e)** Median house values against median income.
(i) For Alameda:
```{r}
ca_pa3 %>% filter(STATEFP==6,COUNTYFP==1) %>% 
  ggplot() +
  geom_point(aes(x = Median_household_income, 
    y = Median_house_value),na.rm = TRUE) +
  labs(x = "Median Household Income",
  y = "Madian House Value",title = "Median House Values vs Median Income
  For Alameda") 
```
(ii) For Santa Clara:
```{r}
ca_pa3 %>% filter(STATEFP==6,COUNTYFP==85) %>% 
  ggplot() +
  geom_point(aes(x = Median_household_income, 
    y = Median_house_value),na.rm = TRUE) +
  labs(x = "Median Household Income",
  y = "Madian House Value",title = "Median House Values vs Median Income
  For Santa Clara") 
```
(iii) For Allegheny:
```{r}
ca_pa3 %>% filter(STATEFP==42,COUNTYFP==3) %>% 
  ggplot() +
  geom_point(aes(x = Median_household_income, 
    y = Median_house_value),na.rm = TRUE) +
  labs(x = "Median Household Income",
  y = "Madian House Value",title = "Median House Values vs Median Income
  For Allegheny") 
```

**MB.Ch1.11. Run the following code:**
```{r}
gender <- factor(c(rep("female", 91), rep("male", 92)))
table(gender)
```
Explain: table() gives the frequency of each element "female" and "male".

```{r}
gender <- factor(gender, levels=c("male", "female"))
table(gender)
```
Explain: the output gives the frequency in levels, that is "male" and "female".

```{r}
gender <- factor(gender, levels=c("Male", "female"))
# Note the mistake: "Male" should be "male"
table(gender)
```
Explain: the output gives the frequency of levels, that is "Male" and "female". While there is no "Male" in gender, so the value of Male is 0.

```{r}
table(gender, exclude=NULL)
```
Explain:The argument "exclude=NULL" require to report NA's, so the result gives <NA>=92 with respond to "male" that is not in the levels.

```{r}
rm(gender)  # Remove gender
```
     
**MB.Ch1.12.** 

```{r}
k<-0
Func <- function(x,cutoff_value){
  for(i in x){
    if (i>cutoff_value){
      k<-k+1
    }
  }
  prop <- k/length(x)
  return(prop)
}
```


(a) For the sequence of numbers 1, 2, . . . , 100, we set the value cutoff 35,67.3 and 89, then the expected proportion result is 0.65, 0.33 and 0.11 correspondingly. And following codes check this.
```{r}
x<-1:100
vc<-35
Func(x,vc)
```
```{r}
vc<-67.3
Func(x,vc)
```
```{r}
vc<-89
Func(x,vc)
```

**MB.Ch1.18. **
Using following commands, we can convert Rabbit to the required form. 
```{r}
Dose <- unstack(Rabbit, Dose ~ Animal)[,1]
Treatment <- unstack(Rabbit, Treatment ~ Animal)[,1]
BPchange <- unstack(Rabbit, BPchange ~ Animal)
Rabbit.df <- data.frame(Treatment, Dose, BPchange)
Rabbit.df
```

