---
title: "Homework 1"
author: "Shen Dingtao  3170104764"
output: pdf_document
---
```{r setup, message = F, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(DAAG)
```

**1. Solution**:

**(a)** First, we load the Iowa data set into R and make it a data frame called `iowa.df` with following command.
```{r}
iowa.df<-read.csv("data/iowa.csv",header=T,sep=";")
```

**(b)** Use the following command to show the dim of `iowa.df`:
```{r}
dim(iowa.df)
```
The result shows that `iowa.df` has 33 rows and 10 columns.

**(c)** Use the following command to show the names of the columns of `iowa.df`:
```{r}
colnames(iowa.df)
```

**(d)** With the following command, we can get the value of row 5, column 7 of `iowa.df` directly:
```{r}
iowa.df[5,7]
```

**(e)** Use the following command to display the second row of `iowa.df` in its entirety:
```{r}
iowa.df[2,]
```
**2. Syntax and class-typing.**

**Solution:**

**(a)** First, We try these following commands in console and get the results:

```{r}
vector1 <- c("5", "12", "7", "32")
max(vector1)
sort(vector1)
```
```
sum(vector1)
```
And the fourth command is error.

Explain: The first command `vector1 <- c("5", "12", "7", "32")` creates a vector of character type, not integer type. So when the parameter of `max()` and `sort()` is `vector1`, the objects to be compared are characters, that is "12" < "32" < "5" < "7", as above result shows. In this case, the function `sum()` can't be executed on character type, so the fourth command is an error.

**(b)** 

1) Assigns a character and two integers to a vector, coercing it into a vector of characters.

```{r}
(vector2 <- c("5",7,12))
```
Thus, `vector2[2]="7", vector2[3]="12"`, they are not numeric, which means `vector2[2]+vector2[3]` is an error. 

2) The function `data.frame` create data frames:
```{r}
dataframe3 <- data.frame(z1="5",z2=7,z3=12)
dataframe3
```
So the first row gives the value of variables `z1,z2,z3`, and `dataframe3[1,2]=7,dataframe3[1,3]=12`, thus
```{r}
dataframe3[1,2]+dataframe3[1,3]
```

3) The function `list()` returns a list or dotted pair list composed of its arguments with each value either tagged or untagged. Thus,
```{r}
list4 <- list(z1="6", z2=42, z3="49", z4=126)
list4
```
And to access elements of the list, `[[ ]]` drops the names and structures, `[ ]` doesn't. That is
```{r}
list4[[2]]
list4[[4]]
```
They are numeric. Adds the two integers together.
```{r}
list4[2]
list4[4]
```
Adds two lists of length 1 together, which is an error.
```{r}
list4[[2]]+list4[[4]]
```
Adds the two integers together.

**3. Working with functions and operators.**

**Solution:**

(a) Use following command to create the sequence of numbers from 1 to 10000 in increments of 372.
```{r}
seq1=seq(1,10000,by = 372)
seq1
```
Use the following command to create a sequence between 1 and 10000 that is exactly 50 numbers in length.
```{r}
seq2=seq(1,10000,length.out=50)
seq2
```

(b) `rep(1:3, times=3)` repeats the whole vector `1:3` for three times, that is
```{r}
rep(1:3, times=3)
``` 
While `rep(1:3, each=3)` repeats each element of the vector `1:3` for 3 times. That is 
```{r}
rep(1:3, each=3)
```
**MB.Ch1.2.** 
Create a new data frame `part_orings` by extracting these rows from orings
```{r}
part_orings<-orings[c(1,2,4,11,13,18),]
```
```{r}
part_orings
```
Plot total incidents against temperature for this new data frame:
```{r}
ggplot(data = part_orings) +
     geom_point(aes(x = Temperature, y = Total))
```
Plot total incidents against temperature for the full data set:
```{r}
ggplot(data = orings) +
     geom_point(aes(x = Temperature, y = Total))
```

**MB.Ch1.4. For the data frame ais (DAAG package)**

**(a)** Use the function str() to get information on each of the columns:
```{r}
str(ais)
```
Use `complete.case()` to determine the rows in which one or more values is missing.
```{r}
complete.cases(ais)
```
The result shows that there is no missing value.

**(b)** There are two methods to make a table that shows the numbers of males and females for each different sport.
Solution 1:
```{r}
new_ais1<-table(ais[,13],ais[,12])
```
```{r}
new_ais1
```
Solution 2:
```{r}
new_ais2 <- ais %>% 
  group_by(sport) %>% 
  count(sex)
(new_ais<-spread(new_ais2,sex,n))
```
To determine if there is a large imbalance (e.g., by a factor of more than 2:1 or less than 1:2) in the numbers of the two sexes, we add a column `fac` to show the factor `f/m`:
```{r}
new_ais %>% mutate(fac=f/m)
```
The result implies that there is a large imbalance in Gym, Netball, in which female dominates(`f/m` > 2). While in T_Sprnt and W_Polo, male dominates(`f/m` < 0.5)  

**MB.Ch1.6.**
```{r}
Manitoba.lakes <- data.frame(c("Winnipeg","Winnipegosis","Manitoba",
                               "SouthernIndian","Cedar","Island",
                               "Gods","Cross","Playgreen"),
                             elevation=c(217,254,248,254,253,227,178,207,217),
                             area=c(24387,5374,4624,2247,1353,1223,1151,755,657))
```
Assign the names of the lakes using the row.names()
function:
```{r}
row.names(Manitoba.lakes)<-Manitoba.lakes[,1]
```
```{r}
rownames(Manitoba.lakes)
```
```{r}

```
**(a)** Use the following code to plot log2(area) versus elevation, adding labeling information (there is an extreme value of area that makes a logarithmic scale pretty much essential):
```{r}
attach(Manitoba.lakes)
plot(log2(area) ~ elevation, pch=16, xlim=c(170,280))
# NB: Doubling the area increases log2(area) by 1.0
text(log2(area) ~ elevation, labels=row.names(Manitoba.lakes), pos=4)
text(log2(area) ~ elevation, labels=area, pos=2) 
title("Manitoba’s Largest Lakes")
```
Captions: the labeling on y-axis gives a logarithmic scale that shows the logs base 2 of area. And the labeling on the points gives the corresponding name of the lake and the actual area of the lake. In this way, the scale on the y-axis will increase by one unit, that is 1, while the area is doubled. 

(b) Repeat the plot and associated labeling, now plotting area versus elevation, but specifying log="y" in order to obtain a logarithmic y-scale.
```{r}
plot(area ~ elevation, pch=16, xlim=c(170,280), log="y")
text(area ~ elevation, labels=row.names(Manitoba.lakes), pos=4, ylog=T)
text(area ~ elevation, labels=area, pos=2, ylog=T) 
title("Manitoba’s Largest Lakes")

```
**MB.Ch1.7.** 
**(a)**
```{r}
dotchart(area,labels=Manitoba.lakes[,1],xlab="area")
```
**(b)**
```{r}
dotchart(log2(area),labels=Manitoba.lakes[,1],xlab="log2(area)")
```

**MB.Ch1.8.** The lower bound for the area of Manitoba covered by water is just the whole area of all lakes in Manitoba. That is
```{r}
sum(Manitoba.lakes$area)
```

